---
title: "Data Wrangling"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Module 2

## Load in the FAS litters data, and clean/rename

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

```


Loading,cleaning and categorizing pups data frame, output 3 rows only :
```{r}
options(tibble.print_min = 3)
pups_df = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_df = janitor::clean_names(pups_df)
```

## looking at Dplyr functions

## function "Select"

Choose some columns and not others.

  1. Specify columns to keep:
```{r}
select(litters_df, group, litter_number, gd0_weight, pups_born_alive)
```

  2. Specify range of columns
```{r}
select(litters_df, group:gd_of_birth)
```
  
  3. Select by specifying columns to remove
```{r}
select(litters_df, -pups_survive)
```

  4. Renaming columns ...
```{r}
select(litters_df, GROUP = group, LITTer_NumBer = litter_number)
```
  
  or
```{r}
rename(litters_df, GROUP = group, LITTer_NumBer = litter_number)
```
  
  
"Select" helpers: starts_with, ends_with, contains

```{r}
select(litters_df, starts_with("gd"))
```

### Reorganizing columns with "everything()", or "relocate"
Select litter_number, pups_survive as 1st two columns, keep everything
```{r}
select(litters_df, litter_number, pups_survive, everything())
```
Relocate does the same thing:
```{r}
relocate(litters_df, litter_number, pups_survive)
```


## function "filter"

Filter gestational-day-0 weight < 22
```{r}
filter(litters_df, gd0_weight < 22)
```

Filter gd_of_birth equal to 20
```{r} 
filter(litters_df, gd_of_birth == 20)
```

Filter gd_of_birth NOT equal to 20
```{r}
filter(litters_df, !(gd_of_birth == 20))
```


### Learning assessment
1. selecting litter number, sex, and PD ears in "pups" data:

```{r}
select(pups_df, litter_number, sex, pd_ears)
```

2a."Filter" to include only pups with sex 1.
```{r}
filter(pups_df, sex == 1)

```

2b. "Filter" pups with PD walk less than 11 and sex 2.
```{r}
filter(pups_df, pd_walk < 11, sex == 2)
```
#### End assessments 1, 2


## Function "mutate" 

Create new variable "wt_gain" using "mutate"
```{r}
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight)
```

Mutate by creating new variable, and modifying  existing variable:
```{r}
mutate(litters_df,
  wt_gain = gd18_weight - gd0_weight,
  group = str_to_lower(group))
```


## Function "Arrange" 
Arrange data by pups_born_alive (will be in increasing order)
```{r}
arrange(litters_df, pups_born_alive)
```

Arrange by two variables: pups_born_alive and gd0_weight. (gd0_weight will be arranged within pups_born_alive)

```{r}
arrange(litters_df, pups_born_alive, gd0_weight)
```


## "%>%" (pipe operation)
Read litters dataframe -> clean names -> select all columns except pups_survive -> mutate df by adding variable wet_gain -> drop rows with missing values from column gd0_weight:

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  select(-pups_survive) %>% 
  mutate(wt_gain = gd18_weight - gd0_weight) %>% 
  drop_na(gd0_weight)
  
```




#### Learning assessment 
3a. Create a variable that subtracts 7 from PD pivot in data "pups"

```{r}
mutate(pups_df,
  pivot_minus7 = pd_pivot - 7)
```

3b. Create a variable that is the sum of all the PD-variables

```{r}
mutate(pups_df, 
       pd_sum = pd_ears + pd_eyes + pd_pivot + pd_walk)
```

4. Piping Assessment:
load the pups data, 
clean the variable names, 
filter the data to include only pups with sex 1, 
remove the PD ears variable, 
create a variable that indicates whether PD pivot is 7 or more days.
```{r}
read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>% 
  filter(sex == 1) %>% 
  select(-pd_ears) %>% 
  mutate(pd_pivot_gt7 = pd_pivot > 7)
```

#### End assessment 3, 4

End module 2



