---
title: "Data Wrangling"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Module 1 
Read in "Litters" dataset, using Relative path, and
Change column names to lower snake case

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)

```

## Take a look at the data, different ways.

Printing in the console.

```{r}
litters_df
```

Checking head of dataset.
```{r}
head(litters_df)
```

Checking tail of dataset.
```{r}
tail(litters_df)
```

```{r}
skimr::skim(litters_df)
```


## options to read_csv

```{r skip 10 rows, first row is not column names}
litters_df = read_csv("./data/FAS_litters.csv", skip = 10, col_names = FALSE)
```

check out `?read_csv()` for more information. 


## Parse data with column specification:

```{r}
cols(
Group = col_character(),
 `Litter Number` = col_character(),
 `GD0 weight` = col_double(),
 `GD18 weight` = col_double(),
 `GD of Birth` = col_double(),
 `Pups born alive` = col_double(),
 `Pups dead @ birth` = col_double(),
 `Pups survive` = col_double()
)
```



### Learning assessment ----------------------
#### Assessment 1:Import and load data "pups"

```{r}
pups_df = read_csv("./data/FAS_pups.csv")
pups_df = janitor::clean_names(pups_df)

```

#### Assessment 2: Parsing data Pups with column specification:

```{r}
pups_df = read_csv(file = "./data/FAS_pups.csv",
  col_types = cols(
    `Litter Number` = col_character(),
    Sex = col_integer(),
    `PD ears` = col_integer(),
    `PD eyes` = col_integer(),
    `PD pivot` = col_integer(),
    `PD walk` = col_integer()
  )
)
skimr::skim(pups_df)
```
### End Assessments 1, 2 ----------------


## readr (read_csv) versus base r fuction (read.csv)
#### assessment #3: 

```{r}
pups_base = read.csv("./data/FAS_pups.csv")
pups_readr = read_csv("./data/FAS_pups.csv")

pups_base
pups_readr
```


Accessing sex variable

```{r}
pups_base$S
pups_readr$S 
```
Did not work in readr! readr needs a full variable name

### End assessment 3 

## End Module 1 


