---
title: "Data Wrangling"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Module 2
Read data, print only 3 lines for tibble, and
Change column names.

```{r}
options(tibble.print_min = 3)

litters_data = read_csv("./data/FAS_litters.csv",
  col_types = "ccddiiii")
litters_data = janitor::clean_names(litters_data)

pups_data = read_csv("./data/FAS_pups.csv",
  col_types = "ciiiii")
pups_data = janitor::clean_names(pups_data)
```


### function "Select"

Select columns from "Litters" dataset:
  1. Specify names to select
```{r}
select(litters_data, group, litter_number, gd0_weight, pups_born_alive)
```

  2. Specify range to select
```{r}
select(litters_data, group:gd_of_birth)
```
  
  3. Select by specifying columns to remove
```{r}
select(litters_data, -pups_survive)
```


### Reorganizing columns with "everything()", or "relocate"

```{r}
select(litters_data, litter_number, pups_survive, everything())
```

```{r}
relocate(litters_data, litter_number, pups_survive)
```


### Learning assessment
1. selecting litter number, sex, and PD ears in "pups" data:

```{r}
select(pups_data, litter_number, sex, pd_ears)
```

2a."Filter" to include only pups with sex 1.
```{r}
filter(pups_data, sex == 1)

```

2b. "Filter" pups with PD walk less than 11 and sex 2.
```{r}
filter(pups_data, pd_walk < 11, sex == 2)
```
#### End assessments 1, 2


## "mutate" : create new variable "wt_gain" using "mutate"
```{r}
mutate(litters_data,
  wt_gain = gd18_weight - gd0_weight)
```

#### Learing assessment 
3a. Create a variable that subtracts 7 from PD pivot in data "pups"

```{r}
mutate(pups_data,
  pivot_minus7 = pd_pivot - 7)
```

3b. Create a variable that is the sum of all the PD-variables

```{r}
mutate(pups_data, 
       pd_sum = pd_ears + pd_eyes + pd_pivot + pd_walk)
```

4. Piping Assessment:
loads the pups data, 
clean the variable names, 
filter the data to include only pups with sex 1, 
remove the PD ears variable, 
create a variable that indicates whether PD pivot is 7 or more days.
```{r}
read_csv("./data/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>% 
  filter(sex == 1) %>% 
  select(-pd_ears) %>% 
  mutate(pd_pivot_gt7 = pd_pivot > 7)
```

#### End assessment 3, 4

End module 2




