---
title: "tidy data and relational datasets"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Module 3

Limit the number of lines printed in a tibble to 5.

```{r}
options(tibble.print_min = 5)
```

Load "PULSE biomarker"s dataset (using "haven" package since data is not csv, it's in sas).

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_df
```

Data Description: For each ID, data shows 4 different observations for bdi score.

Change data from wide format to long format:

## "Pivot_longer"
Tidy and transform the PULSE dataset into a usable format.

```{r}
pulse_df_tidy = 
  pulse_df %>%
   pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_", 
    values_to = "bdi"
    ) 
```

Rwrite, commbine, and extend(to add a mutuate)

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>% 
   pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_", 
    values_to = "bdi"
   ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```



"Pivot_wider"
Create a table:

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result
```

Rearrange with pivot_wider to write "time" subcategories in columns:

```{r}
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean")
```




