---
title: "tidy data and relational datasets"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Module 3

Limit the number of lines printed in a tibble to 5.

```{r}
options(tibble.print_min = 5)
```

Load "PULSE biomarker"s dataset (using "haven" package since data is not csv, it's in sas).

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()

pulse_df
```

Data Description: For each ID, data shows 4 different observations for bdi score.

Change data from wide format to long format:

## "Pivot_longer"
Tidy and transform the PULSE dataset into a usable format.

```{r}
pulse_df_tidy = 
  pulse_df %>%
   pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_", 
    values_to = "bdi"
    ) 
```

Rwrite, commbine, and extend(to add a mutuate)

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>% 
   pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_", 
    values_to = "bdi"
   ) %>% 
  relocate(id, visit) %>% 
  mutate(visit = recode(visit, "bl" = "00m"))
```


## "Pivot_wider"
Create a (tidy) data table:

```{r}
analysis_result = 
  tibble(
    group = c("treatment", "treatment", "placebo", "placebo"),
    time = c("pre", "post", "pre", "post"),
    mean = c(4, 8, 3.5, 4)
)

analysis_result
```

Rearrange with pivot_wider to write "time" subcategories in columns:

```{r}
pivot_wider(
  analysis_result, 
  names_from = "time",   ## column names are from "time"
  values_from = "mean"
  )
```
Easier to read after pivot_wider!


## Binding rows

Using the LotR data.

First step: import each table.

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>% 
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>% 
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>% 
  mutate(movie = "return_king")
```

Bind all the rows together

```{r}
lotr_tidy =
  bind_rows(fellowship_ring, two_towers, return_king) %>% 
  janitor::clean_names() %>% 
  relocate(movie) %>% 
  pivot_longer(
    female:male,
    names_to = "gender",
    values_to = "words"
  )
```

## Joining datasets

Import and clean the FAS datasets.

```{r}
pups_df =
  read_csv("./data/FAS_pups.csv") %>% 
  janitor::clean_names() %>% 
  mutate(sex = recode(sex, `1` = "male", `2` = "female"))  # recoding sex column from numbers to labels.

litters_df = 
  read_csv("./data/FAS_litters.csv") %>% 
  janitor::clean_names() %>% 
  relocate(litter_number) %>%    # relocate this column to beginning
  separate(group, into = c("dose", "day_of_tx"), sep = 3)  
#separate groups column in to two, have to specify where to split the data value: separate after 3 characters!
```

Next up, time to join them!
Call it fas_df = fetal alcohol syndrome dataframe.
Join litter df into pups df, using litter_number as Key.

```{r}
fas_df = 
  left_join(pups_df, litters_df, by = "litter_number") %>%
  arrange(litter_number) %>% 
  relocate(litter_number, dose, day_of_tx)
```

